N-dimensional Tic Tac Toe, and Adventure in Modules

### 2017-05-12 paper/paper.tex
IO section done

	\section{Design Decisions}
  % (Most important!) Describe in detail 3-5 design decisions you made during your
  % project, with respect to abstraction and/or modularity. For example, what is a
  % piece of information that you chose to hide and why? How did you decide on a
  % particular way of separating your system into modules compared to another. How
  % did you reduce coupling between two modules? These design decisions should
  % emphasize how you applied ideas from the papers we’ve read (and associated
  % discussions) to the design of your project. It might be helpful here to
  % include snippets from old versions (or hypothetical alternative versions) of
  % your project to compare and contrast with the final version.
  \subsection{Creating an Abstract Game Engine}
  \subsection{Higher Ordered Signatures, and the ``Include'' incantation}
  \subsection{Separation of IO, or How I learned to not fight SML in search of Purity}
  \subsection{You can do it in 2-dimensions, but can you do it in n-dimensions!}
  \subsection{The Functor is love, the Functor is life}

>>>>

	\section{Design Decisions}
  % (Most important!) Describe in detail 3-5 design decisions you made during your
  % project, with respect to abstraction and/or modularity. For example, what is a
  % piece of information that you chose to hide and why? How did you decide on a
  % particular way of separating your system into modules compared to another. How
  % did you reduce coupling between two modules? These design decisions should
  % emphasize how you applied ideas from the papers we’ve read (and associated
  % discussions) to the design of your project. It might be helpful here to
  % include snippets from old versions (or hypothetical alternative versions) of
  % your project to compare and contrast with the final version.
  \subsection{Creating an Abstract Game Engine}
  \subsection{Higher Ordered Signatures, and the ``Include'' incantation}
  \subsection{Separation of IO, or How I learned to not fight SML in search of Purity}
  In the first implementation of our project we simulated Haskell's IO monad by
  creating a ``show'' signature, and an ``IO'' functor, as shown below:

  \begin{verbatim}
    signature SHOW =
    sig
        type a
        val show : a -> string
    end

    signature IO =
    sig
        structure S : SHOW

        val printIO : S.a -> unit
        val read : 'a -> string option
        val say : string -> unit
    end

    functor Io (structure Sh : SHOW) : IO =
    struct

      structure S = Sh

      (* append a new line to a str, this is expensive *)
      fun appendNewLine str = implode $ (explode str) @ [#"\n"]

      fun printIO x = print o appendNewLine o S.show $ x

      (* function to get user input, it doesn't do anything with its argument *)
      fun read _ = TextIO.inputLine TextIO.stdIn

      fun say str = print o appendNewLine $ str

    end
  \end{verbatim}
  We decided against this approach i.e. isolating IO into a separate module,
  because it began to pollute our design's dependency graph with unwanted edges.
  In haskell type class instances exist in an global overloaded name space, that
  takes advantage of haskell's dispatch system. So each data type that is
  implemented for the \texttt{Show} typeclass is able to be printed to StdOut
  \textit{without} carrying around its respective show function. This is not the
  case in SML and it was sorely missed in our implementation. If we were to
  follow the Haskell style of IO (separating impure and pure code explicitly),
  then an IO dependency would be requisite anytime we needed to perform any IO.
  Thus, the IO node would become a dominating node in our dependency graph and
  the import of it into every functor would quickly become unweidly and
  boilerplate. Hence, we chose an SML style design where the structures that are
  mapped upon carry any relevant functions with them to maintain a clean
  decomposition in our design. This type of structure bloat, although
  disgraceful to a haskell programming, seems to be encouraged by SML's module
  system.

  \subsection{You can do it in 2-dimensions, but can you do it in n-dimensions!}
  \subsection{The Functor is love, the Functor is life}
